Moral Foundations of U.S. Political News Organizations

### 2024-11-30 manuscript_version/revision/manuscript_metapsych_revision.Rmd
rearranging folders and updating the manuscript

abstract: |
  Partisan differences and diviseness have become an increasing hot topic in psychological research. Many theories have been proposed to explain these differences and divisions including Moral Foundations Theory. The current research seeks to use a linguistic measure of Moral Foundations, the Moral Foundations Dictionary (MFD), to test the theory in terms of predicted partisan differences. In Study 1, we explore the MFD-Moral Foundations Questionnaire's measurement overlap using a multi-trait multi-method model. These latent trait models were used to inform possible improvement of the MFD. Next, we examined two replication extensions of previous work exploring the differences in political articles on moral differences Through web scraping, we extracted articles from popular partisan news sources' websites, calculated MFD word frequencies, and identified words' respective valences. This process attempts to uncover news outlets' positive or negative endorsements of certain moral dimensions concomitant with a particular ideology. In Study 1, we gathered political articles from four sources. We were unable to reveal significant differences in moral endorsements, but we solidified the method to be employed in further research. In Study 2, we expanded their number of sources to 10 and analyzed articles that pertain to two specific topics: the 2018 confirmation hearings of U.S. Supreme Court Justice Brett Kavanaugh and the partial U.S. Government Shutdown of 2018-2019. Once again, no significant differences in moral endorsements were found. Together with past work, the results shed doubt on the validity of the MFD as a reliable measurement tool.

@@

bibliography      : ["./metapsych_bib.bib"] # add new references

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
bibliography: references.bib # delete this 
---

```{r load_packages, include = FALSE}
library(papaja)
library(MuMIn)
library(reshape)
library(nlme)
library(ggplot2)
library(expss)
library(MOTE)
library(quanteda)
library(quanteda.textstats)
library(tm)
library(ngram)
library(ggplot2)
library(reshape)
library(knitr)
library(performance)
library(dplyr)
library(rio)
library(mice)
library(semPlot)
library(lavaan)

@@

The internal consistency of the MFQ from @Graham2011 was $\alpha$ = .73
averaged across subscales with a range of $\alpha$ = .65-.84. Across six
studies, the MFQ was found to have an average Cronbach's alpha of .63
for *harm*, .64 for *fairness*, .56 for *ingroup*, .59 for *authority*,
and .71 for *purity* [@Federico2013; @Graham2009; @Graham2012;
@Weber2013]. Test-retest reliability was *r* = .68-.82 using a sample of
123 college students. Confirmatory factor analysis supported a
well-fitted five-factor model (*harm/care*, *fairness/reciprocity*,
*ingroup/loyalty*, *authority/respect*, and *purity/sanctity*) over two,
individual (*harm* and *fairness*) versus group (*ingroup*, *authority*,
and *purity*) foundations, or three, autonomy (*harm*, *fairness*),
community (*ingroup*, *authority*), and divinity (*purity*) ethics,
foundations factor model. The five-factor structure also fit for
non-Western samples, thus, providing evidence of the MFQ
generalizability. Convergent validity was supported with correlations on
other measures of morality [@Schwartz; @Graham2011].

@@

In contrast to this approach, @Sagi2013 explored how moral words were
used paired with other co-occurring concepts using Latent Semantic
Analysis (LSA). They examined three different moral issues in different
contexts to piece out specific moral words and their collocates. First,
they looked at how moral words were used in relation to the World Trade
Center compared to the Empire State Building in the New York Times from
1987-2007. After 9-11, the number of moral words associated with the
World Trade Center increased, specifically *harm* words from the MFD.
Second, they considered the changes in how moral words were paired with
mosque used in blogs as a response to the debate of building a mosque
near Ground Zero following 9-11. They found words from the MFD were used
more often with mosque during the main debate and then the co-occurrence
decreased afterwards. Lastly, they examined moral language tied to the
abortion debate in Congress. Republicans used more moral language
overall; more specifically, Republicans tended to use more words
associated with the *purity* foundation; while Democrats used more words
associated with the *fairness* foundation.

@@

# Method

@@

1.  Model 1 *Correlated Traits Correlated Methods*
    a.  Original MFD Words
    b.  Original MFD Words + Study 1.1 + Study 1.2 Words
    c.  Original MFD Words + Study 1.1 Reduced + Study 1.2 Words
    d.  Determine if model suggests a feasible structure (fit indices
        adequate + loadings adequate)
    e.  Pick the best model to continue (using model comparison)
2.  Model 2 *No Traits Correlated Methods*
3.  Model 3 *Perfectly Correlated Traits Correlated Methods*
4.  Model 4 *Correlated Traits Uncorrelated Methods*

```{r fig-mtmm, fig.cap="Representation of the multi-trait multi-method model for the MFD and MFQ. Not all measured variables (squares) are shown for ease of reading. The left hand side represents the traits or foundation areas of MFT, while the right hand side represents the two measurement tools for the MFT. "}
knitr::include_graphics("mtmm_mfd.png")
```

@@

```{r demos-mtmm-study-3.1, include = FALSE}
exp1data <- import("../../study1_data/study1.1_fixed_associations.csv")

@@

# outliers on MFQ only 
mahal1 <- mahalanobis(study1.1DF[ , -c(1:6,22,23)], 
                    colMeans(study1.1DF[ , -c(1:6,22,23)], na.rm = TRUE),
                    cov(study1.1DF[ , -c(1:6,22,23)], use="pairwise.complete.obs"))
#mahal1
cutoff1 <- qchisq(1 - .001, ncol(study1.1DF[ , -c(1:6,22,23)]))
df1 <- ncol(study1.1DF[ , -c(1:6,22,23)])
summary1 <- summary(mahal1 < cutoff1) # nas will be excluded because they don't have the full scale 
noout1.1 <- subset(study1.1DF, mahal1 < cutoff1)

# assumptions
# correlations for MFQ
correlations1 <- cor(noout1.1[,-c(1:6,22,23)], use="pairwise.complete.obs")
correlations1
symnum(correlations1)

# making the random stuff
random1 <- rchisq(nrow(noout1.1), 7)
fake1 <- lm(random1 ~ ., data = noout1.1[ , -c(1:6, 22, 23)])
# normal
standardized1 <- rstudent(fake1)
hist(standardized1, breaks=15) # some skew

# linear
{qqnorm(standardized1) 
abline(0,1)}
# fairly linear

# homog + s
fitvalues1 <- scale(fake1$fitted.values)
{plot(fitvalues1, standardized1) 
abline(0,0)
abline(v = 0)}
# actually pretty good minus that one outlier

@@

`r nrow(exp1data)` participants were assessed from a large Midwestern
university. Participants were given course credit for their introductory
psychology course for completing the study. `r unname(summary1[4])`
participants had missing data on the MFQ and were excluded from
analyses. The MFQ data was screened for multivariate outliers with
Mahalanobis distance as described in @Tabachnick2012, and
`r unname(summary1[2])` outliers were found using
$\chi^2_{p < .001}$(`r df1`) = `r apa_num(cutoff1)`. These participants
were excluded from further analyses, representing `r nrow(noout1.1)`
final participants. Participants were asked to denote their political
party, and `r printnum(party["Democrat"]/sum(party)*100,digits = 1)`%
indicated they were Democrats,
`r printnum(party["Republican"]/sum(party)*100,digits = 1)`% were
Republican, and
`r printnum(party["Independent"]/sum(party)*100,digits = 1)`% indicated
they were Independent. `r printnum(party[4]/sum(party)*100,digits = 1)`%
of values were missing for this question.

@@

A complete example of the survey can be found online at
<https://osf.io/5uqfn>. First, participants were given a description of
associative memory as the relation between words that comes about
through many pairings in writing and speech. Next, the free association
task, similar to that used in @nelson2004 and @dedeyne2019 was described
to the participants as listing the "first word that pops into mind". The
participants were then given three example free association cues,
*lost*, *old*, and *article*. For each cue, participants were asked to
write all the words that come to mind. To elicit free association to the
moral foundation areas, participants were given the following
instructions:

@@

1.  Manual spell checking of free association responses
2.  Word tokenization and stemming
3.  Create a term by document matrix
4.  Remove punctuation and stopwords via the *tm* stopword list
5.  Calculate percent frequency across documents (all words saved
    online)
6.  Remove words with less than 1% mentions across participants

```{r harm-analysis-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q27 <- as.character(noout1.1$Q27)

@@

# subset out to only words with 1% mentions
harm_words <- harm_freq$Word[harm_freq$percent >=1]
# turn sideways to add to dataset
harm_TDM  <- as.data.frame(t(harm_TDM))
harm_TDM  <- harm_TDM[ , harm_words]

@@

```{r fair-analysi-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q29 <- as.character(noout1.1$Q29)

@@

# subset out to only words with 1% mentions
fair_words <- fair_freq$Word[fair_freq$percent >=1]
# turn sideways to add to dataset
fair_TDM <- as.data.frame(t(fair_TDM))
fair_TDM <- fair_TDM[ , fair_words]

@@

```{r ingroup-analysis-3.1, include=FALSE}
#q27 through 34 are the writing columns
noout1.1$Q30 <- as.character(noout1.1$Q30)

@@

# subset out to only words with 1% mentions
ingroup_words <- ingroup_freq$Word[ingroup_freq$percent >=1]
# turn sideways to add to dataset
ingroup_TDM <- as.data.frame(t(ingroup_TDM))
ingroup_TDM <- ingroup_TDM[ , ingroup_words]

@@

```{r authority-analysis-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q32 <- as.character(noout1.1$Q32)

@@

# subset out to only words with 1% mentions
authority_words <- authority_freq$Word[authority_freq$percent >=1]
# turn sideways to add to dataset
authority_TDM <- as.data.frame(t(authority_TDM))
authority_TDM <- authority_TDM[ , authority_words]

@@

```{r purity-analysis-3.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q34 <- as.character(noout1.1$Q34)

@@

# subset out to only words with 1% mentions
purity_words <- purity_freq$Word[purity_freq$percent >=1]
# turn sideways to add to dataset
purity_TDM <- as.data.frame(t(purity_TDM))
purity_TDM <- purity_TDM[ , purity_words]

@@

this section is about the writing

@@

```{r demos-mtmm-study-3.2, include=FALSE, warning=FALSE}
sample1 <- read.csv("study1_data/study1.2_fixed_sample 1.csv", stringsAsFactors = F)
sample2 <- read.csv("study1_data/study1.2_fixed_sample 2.csv", stringsAsFactors = F)

@@

##merge together the writing samples
#q24, q25, q26, q7 
fullsample$writing <- apply(fullsample[, c("Q24", "Q25", "Q26", "Q7")], 
                           1, function(x) toString(na.omit(x)))

##clear out completely columns we don't need
crapcolumns <- c(1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 51:61)
finalsample <- fullsample[ , -crapcolumns]

##accuracy
summary(finalsample[ , 2:40])

finalsample$Q10 <- factor(finalsample$Q10,
                         levels = 1:2,
                         labels = c("Man", "Woman"))
finalsample$Q11 <- as.numeric(finalsample$Q11)
finalsample$Q12 <- factor(finalsample$Q12,
                         levels = 1:7,
                         labels = c("White", "Black", 
                                    "Hispanic", "Asian",
                                    "Native American", 
                                    "Mixed", "Other"))

gender <- table(finalsample$Q10)/sum(table(finalsample$Q10))*100
race <- table(finalsample$Q12)/sum(table(finalsample$Q12))*100

write.csv(finalsample, 'study1_data/study1.2_finalsample.csv', row.names = FALSE)
newMFD <- import('study1_data/study1.2_finalsample_newMFD.csv')
colnames(newMFD)[1:41] <- colnames(finalsample)[1:41]
newMFD <- newMFD[,c(1,42:55)]
finalsample <- merge(finalsample, newMFD, by = "ResponseId")

@@

for (i in 1:nrow(finalsample)) {

  finalsample$writing[i] <- preprocess(finalsample$writing[i], case="lower", remove.punct=TRUE)
  finalsample$wordcount[i] <- string.summary(finalsample$writing[i])$words

@@

#remove small words people
finalwords <- subset(finalsample, wordcount >=25)

@@

##outliers
mahal2 <- mahalanobis(allcolumns[ , -c(1,46:56)], 
                    colMeans(allcolumns[ , -c(1,46:56)], na.rm = TRUE),
                    cov(allcolumns[ , -c(1,46:56)], use="pairwise.complete.obs"))
#mahal2
cutoff2 <- qchisq(1 - .001,ncol(allcolumns[ , -c(1,46:56)]))
cutoff2
ncol(allcolumns[ , -c(1,46:56)])
summary(mahal2 < cutoff2) #Hmm what to do w/ NAs?? 
noout2 <- subset(allcolumns, mahal2 < cutoff2)

##assumptions on the questions i.e., not Q10 to the end
#correlations
correlations2 <- cor(noout2[,-c(1,46:56)], use="pairwise.complete.obs")
#correlations2
symnum(correlations2)

#the random stuff
random2 <- rchisq(nrow(noout2), 7)
fake2 <- lm(random2~., data=noout2[ , -c(1,46:56)])

#normal
standardized2 <- rstudent(fake2)
hist(standardized2, breaks=15)

#linear
{qqnorm(standardized2)   
abline(0,1)}

#homog + s
fitvalues2 <- scale(fake2$fitted.values)
{plot(fitvalues2, standardized2) 
abline(0,0)
abline(v = 0)}

####create the MFQ subscales####
noout2$harmMFQ <- apply(noout2[ , 2:4], 1, sum)
noout2$fairMFQ <- apply(noout2[ , 5:7], 1, sum)
noout2$ingroupMFQ <- apply(noout2[ , 8:10], 1, sum)
noout2$authorityMFQ <- apply(noout2[ , 11:13], 1, sum)
noout2$purityMFQ <- apply(noout2[ , 14:16], 1, sum)
```

Participants were recruited in two waves as part of a larger
investigation on priming political and religious attitudes. Participants
were recruited via an online research system (SONA) and were given
course credit for their participation. `r nrow(finalsample)`
participants were included in the this study. The study was mostly women
(`r printnum(gender["Woman"], digits = 1)`%) and White
(`r printnum(race["White"], digits = 1)`%) participants with a mix of
minority participants: Black (`r printnum(race["Black"], digits = 1)`%),
(`r printnum(race["Hispanic"], digits = 1)`%),
(`r printnum(race["Asian"], digits = 1)`%), Native American
(`r printnum(race["Native American"], digits = 1)`%), Mixed
(`r printnum(race["Mixed"], digits = 1)`%) and Other
(`r printnum(race["Other"], digits = 1)`%). The average listed age was
`r printnum(mean(finalsample$Q11, na.rm = T))` (*SD* =
`r printnum(sd(finalsample$Q11, na.rm = T))`).

@@

Data was again collected via Qualtrics (<https://osf.io/qr3gv>). Four
fake new stories were presented to participants, which were roughly 400
words each. First, all news stories included a few sentences describing
the use of chemical weapons in the Syrian civil war. The news stories
were manipulated with political (Republican v. Democrat) and religious
(religious v. not) quotes in a 2 x 2 design. News stories can be found
in the online materials. Participants also completed the 30-item version
of the MFQ as described previously. In addition to basic demographics
(gender, age), participant political orientation was assessed with the
same scale described in Study 3.1.

After consenting to participate in the study, participants were randomly
shown one of the four new articles about Syria's use of chemical
weapons. Participants were then asked to write for 5-10 minutes about
their reaction to Syria's use of chemical weapons and the needed
response from the United States. The second wave of data collection
included different writing prompts designed to capture more of the moral
foundation areas in their writing. The following writing prompt was
used, "Please write about your attitudes on abortion (or same-sex
marriage or environmentalism) as well as your reason for this stance."
The three prompts were chosen to create a more varied word set by using
topics that should elicit words from each moral foundations category by
soliciting a moral response. In each wave of data collection (Syria
prompts, moral prompts) the prompt material was randomized between
participants. Participants then completed the MFQ, demographics, and the
political orientation scale.

@@

Participant data was first spell checked and screening for inaccurate
responses. Participants who did not write more than fifty words in
response to a given prompt were excluded (*n* =
`r nrow(finalsample) - nrow(finalwords)`). One missing datapoint was
estimated using the *mice* library from *R* [@CITE] for a missing MFQ
question, and all other missing data was present in the demographics
sections, which were not filled in. The MFQ data were screened for
outliers using Mahalanobis distance, and 25 outliers were found at
$\chi^2_{p<.001}$(`r ncol(allcolumns[ , -c(1,32:42)])`) =
`r printnum(cutoff2)`. These data were excluded leading to a final
sample size of `r nrow(noout2)`. Data were screened for assumptions
described in Study 3.1 and were found to be satisfactory.

```{r writing-analysis-3.2, include=FALSE}
##stem the data
for (i in 1:nrow(noout2)) {
  noout2$writing2[i] <- paste(unlist(
    text_tokens(noout2$writing[i], stemmer = "en")), collapse = " ")
}

##create a corpus
overall_corpus <- Corpus(VectorSource(noout2$writing2))
overall_TDM <- as.matrix(TermDocumentMatrix(overall_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

##view the most frequent words
overall_freq <- data.frame(Word = rownames(overall_TDM),
                          Freq = rowSums(overall_TDM),
                          row.names = NULL)
overall_freq$Word <- as.character(overall_freq$Word)
overall_freq$percent <- overall_freq$Freq/nrow(noout2)*100

write.csv(overall_freq, "study1_data/study1.2_freq.csv", row.names = F)

##subset out to only words with 1% mentions
##this goes over 100% but mostly want to subset out things that were only mentioned
##a few times by participants
all_words <- overall_freq$Word[overall_freq$percent >=1]

@@

##final dataset
overall_final <- cbind(noout2[ , c("ResponseId", "harmMFQ", "fairMFQ",
                                  "authorityMFQ", "purityMFQ","ingroupMFQ")],
                   overall_TDM)
```

>>>>

abstract: |
  Partisan differences and diviseness have become an increasing hot topic in psychological research. Many theories have been proposed to explain these differences and divisions including Moral Foundations Theory. The current research seeks to use a linguistic measure of Moral Foundations, the Moral Foundations Dictionary (MFD), to test the theory in terms of predicted partisan differences. In Study 1, we explore the MFD-Moral Foundations Questionnaire's measurement overlap using a multi-trait multi-method model. These latent trait models were used to inform possible improvement of the MFD. Next, we examined two replication extensions of previous work exploring the differences in political articles on moral differences Through web scraping, we extracted articles from popular partisan news sources' websites, calculated MFD word frequencies, and identified words' respective valences. This process attempts to uncover news outlets' positive or negative endorsements of certain moral dimensions concomitant with a particular ideology. In Study 2, we gathered political articles from four sources. We were unable to reveal significant differences in moral endorsements, but we solidified the method to be employed in further research. In Study 3, we expanded their number of sources to 10 and analyzed articles that pertain to two specific topics: the 2018 confirmation hearings of U.S. Supreme Court Justice Brett Kavanaugh and the partial U.S. Government Shutdown of 2018-2019. Once again, no significant differences in moral endorsements were found. Together with past work, the results shed doubt on the validity of the MFD as a reliable measurement tool.

@@

# bibliography      : ["./metapsych_bib.bib"] # add new references

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
bibliography: references.bib 
appendix          :
  - "appendix.Rmd"
---

```{r load_packages, include = FALSE}
library(papaja)
library(MuMIn)
library(reshape)
library(nlme)
library(ggplot2)
library(expss)
library(MOTE)
library(quanteda)
library(quanteda.textstats)
library(tm)
library(ngram)
library(ggplot2)
library(reshape)
library(knitr)
library(performance)
library(dplyr)
library(rio)
library(mice)
library(semPlot)
library(lavaan)
library(broom)
library(tidyr)

@@

The internal consistency of the MFQ from @Graham2011 was $\alpha$ = .73
averaged across subscales with a range of $\alpha$ = .65-.84. Across six
studies, the MFQ was found to have an average Cronbach's alpha of .63
for *harm*, .64 for *fairness*, .56 for *ingroup*, .59 for *authority*,
and .71 for *purity* [@Federico2013; @Graham2009; @Graham2012;
@Weber2013]. Test-retest reliability was *r* = .68-.82 using a sample of
123 college students. Confirmatory factor analysis supported a
well-fitted five-factor model (*harm/care*, *fairness/reciprocity*,
*ingroup/loyalty*, *authority/respect*, and *purity/sanctity*) over two,
individual (*harm* and *fairness*) versus group (*ingroup*, *authority*,
and *purity*) foundations, or three, autonomy (*harm*, *fairness*),
community (*ingroup*, *authority*), and divinity (*purity*) ethics,
foundations factor model. The five-factor structure also fit for
non-Western samples, thus, providing evidence of the MFQ
generalizability. Convergent validity was supported with correlations on
other measures of morality [@Graham2011].

@@

In contrast to this approach, @Sagi2014 explored how moral words were
used paired with other co-occurring concepts using Latent Semantic
Analysis (LSA). They examined three different moral issues in different
contexts to piece out specific moral words and their collocates. First,
they looked at how moral words were used in relation to the World Trade
Center compared to the Empire State Building in the New York Times from
1987-2007. After 9-11, the number of moral words associated with the
World Trade Center increased, specifically *harm* words from the MFD.
Second, they considered the changes in how moral words were paired with
mosque used in blogs as a response to the debate of building a mosque
near Ground Zero following 9-11. They found words from the MFD were used
more often with mosque during the main debate and then the co-occurrence
decreased afterwards. Lastly, they examined moral language tied to the
abortion debate in Congress. Republicans used more moral language
overall; more specifically, Republicans tended to use more words
associated with the *purity* foundation; while Democrats used more words
associated with the *fairness* foundation.

@@

@@

@@

```{r demos-mtmm-study-1.1, include = FALSE}
exp1data <- import("../../study1_data/study1.1_fixed_associations.csv")

@@

# outliers on MFQ only 
mahal1.1 <- mahalanobis(study1.1DF[ , -c(1:6,22,23)], 
                    colMeans(study1.1DF[ , -c(1:6,22,23)], na.rm = TRUE),
                    cov(study1.1DF[ , -c(1:6,22,23)], use = "pairwise.complete.obs"))
#mahal1.1
cutoff1.1 <- qchisq(1 - .001, ncol(study1.1DF[ , -c(1:6,22,23)]))
df1.1 <- ncol(study1.1DF[ , -c(1:6,22,23)])
summary1.1 <- summary(mahal1.1 < cutoff1.1) # nas will be excluded because they don't have the full scale 
noout1.1 <- subset(study1.1DF, mahal1.1 < cutoff1.1)

# assumptions
# correlations for MFQ
correlations1.1 <- cor(noout1.1[,-c(1:6,22,23)], use = "pairwise.complete.obs")
correlations1.1
symnum(correlations1.1)

# making the random stuff
random1.1 <- rchisq(nrow(noout1.1), 7)
fake1.1 <- lm(random1.1 ~ ., data = noout1.1[ , -c(1:6, 22, 23)])
# normal
standardized1.1 <- rstudent(fake1.1)
hist(standardized1.1, breaks = 15) # some skew

# linear
{qqnorm(standardized1.1) 
abline(0,1)}
# fairly linear

# homog + s
fitvalues1.1 <- scale(fake1.1$fitted.values)
{plot(fitvalues1.1, standardized1.1) 
abline(0,0)
abline(v = 0)}
# actually pretty good minus that one outlier

@@

`r nrow(exp1data)` participants were assessed from a large Midwestern
university. Participants were given course credit for their introductory
psychology course for completing the study. `r unname(summary1.1[4])`
participants had missing data on the MFQ and were excluded from
analyses. The MFQ data was screened for multivariate outliers with
Mahalanobis distance as described in @Tabachnick2012, and
`r unname(summary1.1[2])` outliers were found using
$\chi^2_{p < .001}$(`r df1.1`) = `r apa_num(cutoff1.1)`. These participants
were excluded from further analyses, representing `r nrow(noout1.1)`
final participants. Participants were asked to denote their political
party, and `r printnum(party["Democrat"]/sum(party)*100,digits = 1)`%
indicated they were Democrats,
`r printnum(party["Republican"]/sum(party)*100,digits = 1)`% were
Republican, and
`r printnum(party["Independent"]/sum(party)*100,digits = 1)`% indicated
they were Independent. `r printnum(party[4]/sum(party)*100,digits = 1)`%
of values were missing for this question.

@@

A complete example of the survey can be found online at
<https://osf.io/6mv2g>. First, participants were given a description of
associative memory as the relation between words that comes about
through many pairings in writing and speech. Next, the free association
task, similar to that used in @nelson2004 and @dedeyne2019 was described
to the participants as listing the "first word that pops into mind". The
participants were then given three example free association cues,
*lost*, *old*, and *article*. For each cue, participants were asked to
write all the words that come to mind. To elicit free association to the
moral foundation areas, participants were given the following
instructions:

@@

1.  Manual spell checking of free association responses.
2.  Word tokenization and stemming.
3.  Create a term by document matrix.
4.  Remove punctuation and stopwords via the *tm* stopword list.
5.  Calculate percent frequency across documents.
6.  Remove words with less than 1% mentions across participants.

```{r harm-analysis-1.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q27 <- as.character(noout1.1$Q27)

@@

# subset out to only words with 1% mentions
harm_words <- harm_freq$Word[harm_freq$percent >= 1]
# turn sideways to add to dataset
harm_TDM  <- as.data.frame(t(harm_TDM))
harm_TDM  <- harm_TDM[ , harm_words]

@@

```{r fair-analysi-1.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q29 <- as.character(noout1.1$Q29)

@@

# subset out to only words with 1% mentions
fair_words <- fair_freq$Word[fair_freq$percent >= 1]
# turn sideways to add to dataset
fair_TDM <- as.data.frame(t(fair_TDM))
fair_TDM <- fair_TDM[ , fair_words]

@@

```{r ingroup-analysis-1.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q30 <- as.character(noout1.1$Q30)

@@

# subset out to only words with 1% mentions
ingroup_words <- ingroup_freq$Word[ingroup_freq$percent >= 1]
# turn sideways to add to dataset
ingroup_TDM <- as.data.frame(t(ingroup_TDM))
ingroup_TDM <- ingroup_TDM[ , ingroup_words]

@@

```{r authority-analysis-1.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q32 <- as.character(noout1.1$Q32)

@@

# subset out to only words with 1% mentions
authority_words <- authority_freq$Word[authority_freq$percent >= 1]
# turn sideways to add to dataset
authority_TDM <- as.data.frame(t(authority_TDM))
authority_TDM <- authority_TDM[ , authority_words]

@@

```{r purity-analysis-1.1, include = FALSE}
#q27 through 34 are the writing columns
noout1.1$Q34 <- as.character(noout1.1$Q34)

@@

# subset out to only words with 1% mentions
purity_words <- purity_freq$Word[purity_freq$percent >= 1]
# turn sideways to add to dataset
purity_TDM <- as.data.frame(t(purity_TDM))
purity_TDM <- purity_TDM[ , purity_words]

@@

In this study, participants were asked to write about chemical warfare
(wave 1), abortion, same-sex marriage, and environmentalism (wave 2).
These prompts were designed to elicit writing about moral concepts that
would relate to the moral foundations areas for expansion of the
dictionary.

@@

```{r demos-mtmm-study-1.2, include = FALSE, warning = FALSE}
sample1 <- read.csv("../../study1_data/study1.2_fixed_sample 1.csv", stringsAsFactors = F)
sample2 <- read.csv("../../study1_data/study1.2_fixed_sample 2.csv", stringsAsFactors = F)

@@

##merge together the writing samples
#q24, q25, q26, q7 
study1.2DF$writing <- apply(study1.2DF[, c("Q24", "Q25", "Q26", "Q7")], 
                           1, function(x) toString(na.omit(x)))

##clear out completely columns we don't need
crapcolumns <- c(1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 51:61)
study1.2DF <- study1.2DF[ , -crapcolumns]

##accuracy
summary(study1.2DF[ , 2:40])

study1.2DF$Q10 <- factor(study1.2DF$Q10,
                         levels = 1:2,
                         labels = c("Man", "Woman"))
study1.2DF$Q11 <- as.numeric(study1.2DF$Q11)
study1.2DF$Q12 <- factor(study1.2DF$Q12,
                         levels = 1:7,
                         labels = c("White", "Black", 
                                    "Hispanic", "Asian",
                                    "Native American", 
                                    "Mixed", "Other"))

gender <- table(study1.2DF$Q10)/sum(table(study1.2DF$Q10))*100
race <- table(study1.2DF$Q12)/sum(table(study1.2DF$Q12))*100

write.csv(study1.2DF, '../../study1_data/study1.2_finalsample.csv', row.names = FALSE)
newMFD <- import('../../study1_data/study1.2_finalsample_newMFD.csv')
colnames(newMFD)[1:41] <- colnames(study1.2DF)[1:41]
newMFD <- newMFD[,c(1,42:55)]
study1.2DF <- merge(study1.2DF, newMFD, by = "ResponseId")

@@

for (i in 1:nrow(study1.2DF)) {

  study1.2DF$writing[i] <- preprocess(study1.2DF$writing[i], case = "lower", remove.punct = TRUE)
  study1.2DF$wordcount[i] <- string.summary(study1.2DF$writing[i])$words

@@

#remove small words people
finalwords <- subset(study1.2DF, wordcount >= 25)

@@

##outliers
mahal1.2 <- mahalanobis(allcolumns[ , -c(1,46:56)], 
                    colMeans(allcolumns[ , -c(1,46:56)], na.rm = TRUE),
                    cov(allcolumns[ , -c(1,46:56)], use = "pairwise.complete.obs"))
#mahal1.2
cutoff1.2 <- qchisq(1 - .001,ncol(allcolumns[ , -c(1,46:56)]))
cutoff1.2
ncol(allcolumns[ , -c(1,46:56)])
summary(mahal1.2 < cutoff1.2) #Hmm what to do w/ NAs?? 
noout1.2 <- subset(allcolumns, mahal1.2 < cutoff1.2)

##assumptions on the questions i.e., not Q10 to the end
#correlations
correlations1.2 <- cor(noout1.2[,-c(1,46:56)], use = "pairwise.complete.obs")
#correlations1.2
symnum(correlations1.2)

#the random stuff
random1.2 <- rchisq(nrow(noout1.2), 7)
fake1.2 <- lm(random1.2 ~ ., data = noout1.2[ , -c(1,46:56)])

#normal
standardized1.2 <- rstudent(fake1.2)
hist(standardized1.2, breaks = 15)

#linear
{qqnorm(standardized1.2)   
abline(0,1)}

#homog + s
fitvalues1.2 <- scale(fake1.2$fitted.values)
{plot(fitvalues1.2, standardized1.2) 
abline(0,0)
abline(v = 0)}

####create the MFQ subscales####
noout1.2$harmMFQ <- apply(noout1.2[ , 2:4], 1, sum)
noout1.2$fairMFQ <- apply(noout1.2[ , 5:7], 1, sum)
noout1.2$ingroupMFQ <- apply(noout1.2[ , 8:10], 1, sum)
noout1.2$authorityMFQ <- apply(noout1.2[ , 11:13], 1, sum)
noout1.2$purityMFQ <- apply(noout1.2[ , 14:16], 1, sum)
```

Participants were recruited in two waves as part of a larger
investigation on priming political and religious attitudes. Participants
were recruited via an online research system (SONA) and were given
course credit for their participation. `r nrow(study1.2DF)`
participants were included in the this study. The study was mostly women
(`r printnum(gender["Woman"], digits = 1)`%) and White
(`r printnum(race["White"], digits = 1)`%) participants with a mix of
minority participants: Black (`r printnum(race["Black"], digits = 1)`%),
(`r printnum(race["Hispanic"], digits = 1)`%),
(`r printnum(race["Asian"], digits = 1)`%), Native American
(`r printnum(race["Native American"], digits = 1)`%), Mixed
(`r printnum(race["Mixed"], digits = 1)`%) and Other
(`r printnum(race["Other"], digits = 1)`%). The average listed age was
`r printnum(mean(study1.2DF$Q11, na.rm = T))` (*SD*  = 
`r printnum(sd(study1.2DF$Q11, na.rm = T))`).

@@

#### Wave 1

Data was again collected via Qualtrics. Four fake new stories
were presented to participants, which were roughly 400 words each.
First, all news stories included a few sentences describing the use of
chemical weapons in the Syrian civil war. The news stories were
manipulated with political (Republican v. Democrat) and religious
(religious v. not) quotes in a 2 x 2 between-subjects design. News
stories can be found in the online materials. After consenting to
participate in the study, participants were randomly shown one of the
four new articles about Syria's use of chemical weapons. Participants
were then asked to write for 5-10 minutes about their reaction to
Syria's use of chemical weapons and the needed response from the United
States. Participants also completed the 30-item version of the MFQ as
described previously. In addition to basic demographics (gender, age),
participant political orientation was assessed with the same scale
described in Study 1.1.

#### Wave 2

The second wave of data collection included different writing prompts
designed to capture more of the moral foundation areas in their writing.
The following writing prompt(s) were used, "Please write about your
attitudes on abortion (or same-sex marriage or environmentalism) as well
as your reason for this stance." The three prompts were chosen to create
a more varied word set by using topics that should elicit words from
each moral foundations category by soliciting a moral response. The
writing prompts were randomized within participant, and each participant
completed all three prompts. Participants then completed the MFQ,
demographics, and the political orientation scale.

@@

Participant data was first spell checked and screening for inaccurate
responses. Participants who did not write more than fifty words in
response to a given prompt were excluded (*n*  = 
`r nrow(study1.2DF) - nrow(finalwords)`). One missing datapoint was
estimated using the *mice* library from *R* for a missing MFQ question
[@buuren2011], and all other missing data was present in the
demographics sections, which were not filled in. The MFQ data were
screened for outliers using Mahalanobis distance, and `r nrow(finalwords) - nrow(noout1.2)` outliers were
found at $\chi^2_{p<.001}$(`r ncol(allcolumns[ , -c(1,32:42)])`)  = 
`r printnum(cutoff1.2)`. These data were excluded leading to a final
sample size of `r nrow(noout1.2)`. Data were screened for assumptions
described in Study 1.1 and were found to be satisfactory.

```{r writing-analysis-1.2, include = FALSE}
##stem the data
for (i in 1:nrow(noout1.2)) {
  noout1.2$writing2[i] <- paste(unlist(
    text_tokens(noout1.2$writing[i], stemmer = "en")), collapse = " ")
}

##create a corpus
overall_corpus <- Corpus(VectorSource(noout1.2$writing2))
overall_TDM <- as.matrix(TermDocumentMatrix(overall_corpus,
                              control = list(removePunctuation = TRUE,
                                             stopwords = TRUE)))

##view the most frequent words
overall_freq <- data.frame(Word = rownames(overall_TDM),
                          Freq = rowSums(overall_TDM),
                          row.names = NULL)
overall_freq$Word <- as.character(overall_freq$Word)
overall_freq$percent <- overall_freq$Freq/nrow(noout1.2)*100

write.csv(overall_freq, "../../study1_data/study1.2_freq.csv", row.names = F)

##subset out to only words with 1% mentions
##this goes over 100% but mostly want to subset out things that were only mentioned
##a few times by participants
all_words <- overall_freq$Word[overall_freq$percent >= 1]

@@

##final dataset
overall_final <- cbind(noout1.2[ , c("ResponseId", "harmMFQ", "fairMFQ",
                                  "authorityMFQ", "purityMFQ","ingroupMFQ")],
                   overall_TDM)
```

```{r mfq-cor-1.2, include = FALSE}
##harm
harm_cor <- apply(overall_final[ , 7:ncol(overall_final)], 2, function (x) { cor(x, overall_final$harmMFQ)})
